Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-01 examples.tex
+

 \subsubsection{Ensuring Presence of Entry}
 Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) such that we can own twe introduced for the  
In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.

Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
\[
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner\urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$

;;if (!entry->present) {
  ...

;;}
;; uintptr_t next_phys_addr = PTE_PFN_TO_ADDR(entry->pfn);
;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
add rbx, KERNBASE ;; shift physical address by identity mapping constant OFF
$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
\end{lstlisting}
\vspace{-1em}
\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
\label{fig:p2v}
\end{figure}

@@

\begin{figure*}[ht]
\begin{tabular}{p{0.47\textwidth}p{0.47\textwidth}}
  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
;;void pte_initialize(pte_t **table) {
...

;;Allocate a full page for 512 8-byte entries
;;pte_t *local = kalloc();
;;clear the entire table. Ensures present (and reserved, etc.) bits are 0
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp]

;;*table = local;
mov    -0x10[rbp],rax
mov    -0x8[rbp],rdi
mov    rax,[rdi]

;;return;
...

;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) 
push   rbp
mov    rsp,rbp
sub    0x40,rsp
mov    rdi,-0x8[rbp]
mov    rsi,-0x10[rbp]
;;pte_t *pml4_entry = &pml4[PML4EX(va)];
mov    -0x8[rbp],[rsi]
mov    -0x10[rbp],rdi
shr    0x27,rdi
and    0x1ff,rdi
shl    0x3,rdi
add    rdi,rsi
mov    rsi,-0x18[rbp]
;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
mov    -0x18[rbp],rdi
callq  <pte_get_next_table_succ>
mov    rax,-0x20[rbp]
;;pte_t *pdp_entry = &pdp[PDPEX(va)];
...
...
...
;;last call at L1 for virtual address to physical page 
callq  <pte_get_next_table_succ>
mov    rax,-0x40[rbp]
;;return &pt[PTEX(va)];
mov    -0x40[rbp],rax
mov    -0x10[rbp],rsi
shr    0xc,rsi
and    0x1ff,rsi
shl    0x3,rsi
add    rsi,rax
add    0x40,rsp
pop    rbp
retq   
  \end{lstlisting}&
  \begin{lstlisting}[mathescape,basicstyle=\scriptsize\ttfamily]
   $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
;; pte_t *pte_get_next_table_succ(pte_t *entry) {
push   rbp
mov    rsp,rbp
sub    0x20,rsp
mov    rdi,-0x8[rbp]
;;  pte_t *next;
mov    -0x8[rbp],rdi
mov    [rdi],rdi
and    0x1,rdi
mov    rdi,rax
;;  if (!entry->present) {
cmp    0x0,rax
jne    161 <pte_get_next_table_succ+0xa1>
lea    -0x10[rbp],rdi
;;pte_initialize(&next);
callq  70 <pte_initialize>
mov    -0x8[rbp],rdi
mov    [rdi],rax
;;setting the fields of the entry with a fresh allocated page
;;entry->writable = 1
;;entry->user_acc = 1;
;;entry->pfn = PTE_ADDR_TO_PHYS_PFN((uintptr_t) next);
;;entry->present = 1;
and    0xfffffffffffffffd,rax
or     0x2,rax
mov    rax,[rdi]
...
;;}
;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
mov    -0x8[rbp],rax
mov    [rax],rax
shr    0xc,rax
movabs 0xffffffffff,rcx
and    rcx,rax
shl    0xc,rax
mov    rax,-0x18[rbp]
;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
mov    -0x18[rbp],rax
movabs 0x0,rcx	187: R_X86_64_64	KERN_BASE
add    rcx,rax
;;next = (pte_t *) next_virt_addr;
mov    rax,-0x20[rbp]
mov    -0x20[rbp],rax
mov    rax,-0x10[rbp]
;;return next
mov    -0x10[rbp],rax
add    0x20,rsp
pop    rbp
ret
  \end{lstlisting}
  \end{tabular}
\vspace{-1em}
\caption{ in Figure \ref{fig:walkpgdir}.}
\label{fig:getnexttable}
\end{figure*}

\subsubsection{}

>>>>

 \subsubsection{Ensuring Presence of Entry}
 Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.

In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (\textsf{kalloc} -- currently an axiom as it is not part of \textsf{pte} module but the allocator's) to be backing the physical frame address of the entry.
\mytodo{Discuss with Colin: v2p and spec on pteinitialize. After clarification expand this piece a little bit more. Right now we have physical pointsto returned from kalloc so pfn is directly set not through v2p}
Once we make sure that the frame of the page-table entry is backed with an allocated physical page, we can obtain the virtual address of the next level-page table (\textsf{next} of function \textsf{get\_next\_table} in Figure \ref{}) via looking up the \textit{identity mapping} from the physical address (the address value referenced by \textsf{entry->pfn}) to virtual address value to be returned by \textsf{get\_next\_table}.
\begin{figure}\footnotesize
\begin{lstlisting}[mathescape]
;;if (!entry->present) {
  ...

;;}
...
;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
  $\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m) \ast \ulcorner \textsf{OFF} + \paddr > 0 \urcorner \ast \texttt{rbx}\mapsto_{\textsf{r}} \paddr \ast \sumapaces\rtv\delta  \ast \sumpv\paddr\qfrac\true  }_{\rtv}$
add rbx, KERNBASE ;; shift physical address by identity mapping constant OFF
$\specline{\textsf{P} \ast \exists \qfraczero; \qfrac,\textsf{ w, val},  \texttt{rbx}\mapsto_{\textsf{r}} \paddr + \textsf{OFF} \ast \sumapaces\rtv\delta \ast \sumpv\paddr\qfrac\true \ast \vaddr \mapsto_{\texttt{vpte}} \{\textsf{q}, \textsf{qfrac}\} \paddr \;(\texttt{val})}$
$\specline{\ulcorner \vaddr = \paddr + \textsf{OFF} \urcorner \ast \ulcorner \textsf{qfrac} = 1 \leftrightarrow \; \lnot\textsf{entry\_present }(\vale) \urcorner \ast \sumwalkabs\vaddr\qfrac\paddr \ast (R \wand \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)) }_{\rtv}$
\end{lstlisting}
\vspace{-1em}
\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
\label{fig:p2v}
\end{figure}

@@
